---
output: html_document
author: AEN and MCF
title: Neg-tablet Analyses
---

Prelims and libraries.

```{r}
rm(list = ls())
#Load libraries
library(reshape2)
library(dplyr)
library(ggplot2)
library(bootstrap)
library(lme4)
library(knitr)
```
Functions and add some style elements for ggplot2

```{r}
## number of unique subs
n.unique <- function (x) {
  length(unique(x))
}

## for bootstrapping 95% confidence intervals
theta <- function(x,xdata) {mean(xdata[x])}
ci.low <- function(x) {
  quantile(bootstrap(1:length(x), 1000, theta, x)$thetastar, .025)}
ci.high <- function(x) {
  quantile(bootstrap(1:length(x), 1000, theta, x)$thetastar, .975)}

plot.style <- theme_bw() + 
  theme(panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank(), 
        legend.position = "right", 
        axis.line = element_line(colour = "black",size = .5), 
        axis.ticks = element_line(size = .5), 
        axis.title.x = element_text(vjust = -.5), 
        axis.title.y = element_text(angle = 90,vjust = 0.25))
```


## Experiment 1: Adults, between-subjects, referent type & syntactic frame

Load in data

```{r}
d.exp1 <- read.csv("data/exp1.csv")
d.exp1$condition <- factor(d.exp1$condition, 
                           levels = c("noContext", "context"), 
                           labels = c("None", "Target"))
d.exp1$sentence.type <- factor(d.exp1$sentence.type, 
                               levels = c("positive", "negative"), 
                               labels = c("Positive" , "Negative"))

d.exp1$neg.concept <- factor(d.exp1$neg.concept, 
                             levels = c("something", "nothing"),
                             labels = c("Alternative",  "Nonexistence"))

d.exp1$neg.syntax <- factor(d.exp1$neg.syntax, 
                            levels = c("", "has no", "doesn't have"), 
                            labels = c("", "has no X", "doesn't have X"))
```

### Plot data

Histogram of rating scale: 

```{r}
qplot(d.exp1$rating, geom = "histogram") + plot.style
```

Plot mean ratings for every trial type (Table 4 in paper):

```{r}
ms <- d.exp1 %>%
  group_by(subid, condition, sentence.type, truth.value, neg.concept, neg.syntax) %>%
  summarise(subm = mean(rating)) %>%
  group_by(condition, sentence.type, truth.value, neg.concept, neg.syntax) %>%
  summarise(rating = mean(subm),
            high = ci.high(subm),
            low = ci.low(subm))
ms$truth.value <- factor(ms$truth.value, levels=c(TRUE, FALSE), labels=c("True","False"))

kable(ms, digits = 2)

ggplot(data = ms, aes(x = sentence.type, y = rating, 
      color = neg.concept, shape = neg.syntax)) +
  geom_point(stat = "identity", position = position_dodge(width = .5)) + 
  facet_grid(truth.value ~ condition) +
  geom_errorbar(aes(ymin = low, ymax = high), width = 0,
                position = position_dodge(width = .5)) + 
  plot.style
```

Just look at negative sentences:

```{r}
d.exp1.neg <- filter(d.exp1, sentence.type == "Negative")

ms <- d.exp1.neg %>%
  group_by(subid, condition, truth.value, neg.concept, neg.syntax) %>%
  summarise(subm = mean(rating)) %>%
  group_by(condition, truth.value, neg.concept, neg.syntax) %>%
  summarise(rating = mean(subm),
            high = ci.high(subm),
            low = ci.low(subm))
ms$truth.value <- factor(ms$truth.value, levels=c(TRUE, FALSE), labels=c("True","False"))

ggplot(data = ms, aes(x = condition, y = rating, fill = neg.concept)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(truth.value ~ neg.syntax) +
  geom_errorbar(aes(ymin = low, ymax = high), width = 0,
                position = position_dodge(width = .9)) + 
  scale_fill_grey("Negation Type") + 
  xlab("Context") + ylab("Rating") + 
  plot.style
```

Just look at true negatives (Figure 2 in paper)

```{r}
d.exp1.trueneg <- filter(d.exp1, sentence.type ==  "Negative" & truth.value == TRUE)

ms <- d.exp1.trueneg %>%
  group_by(subid, condition, neg.concept, neg.syntax) %>%
  summarise(subm = mean(rating)) %>%
  group_by(condition, neg.concept, neg.syntax) %>%
  summarise(rating = mean(subm),
            high = ci.high(subm),
            low = ci.low(subm))

ggplot(data = ms, aes(x = neg.concept, y = rating, fill = condition)) +
  geom_bar(stat = "identity", position = "dodge") +  
  facet_wrap(~ neg.syntax) + 
  geom_errorbar(aes(ymin = low, ymax = high), width = 0,
                position = position_dodge(width = .9)) + 
  scale_fill_grey("Negation Type") + 
  xlab("Context") + ylab("Rating") + 
  plot.style

```

### Statistical models: 

All data: 
```{r}
#deviation coding
contrasts(d.exp1$sentence.type) <- rbind(.5, -.5)
contrasts(d.exp1$sentence.type)

contrasts(d.exp1$condition) <- rbind(.5, -.5)
contrasts(d.exp1$condition)

contrasts(d.exp1$truth.value) <- rbind(.5, -.5)
contrasts(d.exp1$truth.value)

model.exp1.all <- lmer(rating ~ sentence.type * condition * truth.value +
                         (sentence.type * truth.value | subid) +
                         (sentence.type * truth.value | item), 
                       data = d.exp1)

kable(summary(model.exp1.all)$coefficients, digits = 2)
```

Just negative sentences: 
```{r}
contrasts(d.exp1.neg$condition) <- rbind(.5, -.5)
contrasts(d.exp1.neg$condition)

contrasts(d.exp1.neg$truth.value) <- rbind(.5, -.5)
contrasts(d.exp1.neg$truth.value)

model.exp1.neg <- lmer(rating ~ condition * truth.value +
                         (truth.value | subid) + 
                         (condition * truth.value | item), 
                       data = d.exp1.neg)

kable(summary(model.exp1.neg)$coefficients, digits = 2)
```

Just true negative sentences (Table 1 in paper): 
```{r}
contrasts(d.exp1.trueneg$condition) <- rbind(.5, -.5)
contrasts(d.exp1.trueneg$condition)

contrasts(d.exp1.trueneg$neg.concept) <- rbind(.5, -.5)
contrasts(d.exp1.trueneg$neg.concept)

d.exp1.trueneg$neg.syntax <- factor(d.exp1.trueneg$neg.syntax)
contrasts(d.exp1.trueneg$neg.syntax) <- rbind(.5, -.5)
contrasts(d.exp1.trueneg$neg.syntax)

model.exp1.trueneg <- lmer(rating ~ condition * neg.concept * neg.syntax +
                             (neg.concept * neg.syntax | subid) + 
                             (neg.concept * neg.syntax | item), 
                           data = d.exp1.trueneg)

kable(summary(model.exp1.trueneg)$coefficients, digits = 2)
```

## Experiment 2: Children, between subjects, none vs. target context

Load in data.

```{r}
d.exp2 <- read.csv("data/exp2.csv")

d.exp2$agegroup <- factor(d.exp2$agegroup)

d.exp2$sent.type <- factor(d.exp2$sent.type, 
                           levels = c("positive", "negative"), 
                           labels = c("Positive", "Negative"))

d.exp2$condition <- factor(d.exp2$condition, 
                           labels = c("None", "Target"))

demographics <- read.csv("data/demographics.csv")

d.exp2 <- merge(d.exp2, demographics)

initial.ns <- d.exp2 %>%
  group_by(condition, agegroup) %>%
  summarize(ns = n.unique(subid))
print(initial.ns)
```

Condense scale from 5-point to 3-point

```{r}
d.exp2$resp2 <- 2
d.exp2[d.exp2$resp > 3,]$resp2 <- 3
d.exp2[d.exp2$resp < 3,]$resp2 <- 1
d.exp2$resp2 <- factor(d.exp2$resp2)
```

### Participant exclusions

Exclude subjects who didn't complete at least half of the trials

```{r}
#Reject children who didn't complete at least 8 trials
ntrials <- d.exp2 %>%
  group_by(subid) %>%
  summarize(ntrials = n()) %>%
  filter(ntrials < 8)

for (i in ntrials$subid) {
  d.exp2 <- filter(d.exp2, subid != i)
}

inter.ns <- d.exp2 %>%
  group_by(condition, agegroup) %>%
  summarize(ns = n.unique(subid))
print(inter.ns)
```

Reject subjects who don't understand scale (based on positive sentences):

```{r}
reject <- d.exp2 %>%
  filter(sent.type == "Positive") %>% #Only look at positive sentences
  group_by(subid) %>%
  mutate(total = n()) %>% #get total # of positive sentences child saw
  group_by(subid, condition, truth, total, resp2) %>%
  filter((truth == "True" & resp2 == 3) | (truth == "False" & resp2 == 1)) %>% #Get # "good" for true pos and "bad" for false pos
  summarize(counts = n()) %>% 
  group_by(subid, condition, total) %>%
  summarize(counts = sum(counts)) %>% #total # "correct" responses
  mutate(prop = counts/total) %>% #proportion correct
  filter(prop < .6) #reject kids who got < .6 "correct" (this allows for 2/6 "mistakes")

for (i in reject$subid) {
  d.exp2 <- filter(d.exp2, subid != i)
}
```

Make sure there aren't any kids who just used one side of scale. Reject kids who only chose a single data point. Check final ns & age info.

```{r}
scaleUse <- aggregate(resp2 ~ subid, d.exp2, n.unique)
table(scaleUse$resp2) #Are any resp2 = 1

final.ns <- d.exp2 %>%
  group_by(condition, agegroup) %>%
  summarize(ns = n.unique(subid), 
            ages = mean(age),
            min.age = min(age),
            max.age = max(age))
print(final.ns)

gender <- d.exp2 %>%
  group_by(condition, agegroup, gender) %>%
  summarize(ns = n.unique(subid))
print(gender)
```

### Plot data, all ages

Categorize kids based on response type

```{r}
tn_responses <- d.exp2 %>%
  filter(sent.type == "Negative" & truth == "True") %>%
  group_by(subid) %>%
  mutate(total = n()) %>%
  group_by(subid, condition, total, resp2) %>%
  summarize(counts = n()) %>%
  mutate(prop = counts/total)

category <- dcast(tn_responses, subid + condition ~ resp2)
names(category) <- c("subid","condition","bad","neutral","good")
category[is.na(category)] <- 0

category$type <- "other"
#category[category$neutral > .6,]$type <- "tn_neutral"
category[category$bad > .6,]$type <- "tn_bad"
category[category$good > .6,]$type <- "tn_good"

cat_counts <- category %>%
  group_by(condition, type) %>%
  summarise(counts = n())

cat_counts$type <- factor(cat_counts$type, levels = c("tn_bad","tn_good","other"), labels = c("True Negatives = Bad", "True Negatives = Good", "Inconsistent/Other"))

ggplot(data = cat_counts, aes(x = condition, y = counts, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") + 
  scale_fill_hue("Response Type") +
  ylab("Count") + xlab("Context Condition") +
  plot.style
```

Mean responses to all sentences (Table 5 in paper):

```{r}
ms <- d.exp2 %>%
  group_by(subid, condition, sent.type, truth) %>%
  summarise(subm = mean(resp)) %>%
  group_by(condition, sent.type, truth) %>%
  summarise(m = mean(subm),
            cih = ci.high(subm),
            cil = ci.low(subm))
ms$truth <- factor(ms$truth, levels=c("True","False"))

kable(ms, digits = 2)

ggplot(data = subset(ms), 
      aes(x = truth, y = m, fill =  condition)) + 
  facet_grid(. ~ sent.type) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = cil, ymax = cih), 
                position = position_dodge(.9), width = 0) + 
  scale_fill_grey("") +
  xlab("Context") + ylab("Response") +
  scale_y_continuous(limits = c(0, 5), breaks = seq(1,5,1)) +
  plot.style
```

Just negative trials (Figure 3 in paper):

```{r}
ggplot(data = subset(ms, sent.type == "Negative"), 
      aes(x = truth, y = m, fill = condition)) + 
  #facet_grid(. ~ truth) + 
  geom_bar(stat = "identity", position = "dodge") + 
  geom_errorbar(aes(ymin = cil, ymax = cih), 
                position = position_dodge(.9), width = 0) + 
  scale_fill_grey("") +
  xlab("Context") + ylab("Response") +
  scale_y_continuous(limits = c(0, 5), breaks = seq(1,5,1)) +
  plot.style
```

Histogram of trial ratings (Figure 4 in paper): 

```{r}
truenegs <- filter(ms, truth == "True" & sent.type ==  "Negative")

#make df for histogram (for formatting reasons)
hist_data <- d.exp2 %>%
  filter(truth == "True" & sent.type == "Negative") %>%
  group_by(condition, resp) %>%
  summarise(count = n())

ggplot(data = hist_data, aes(y = count, x = resp, 
      fill = condition)) +
  geom_bar(width = .5, position = position_dodge(.6), stat = "identity") +
  geom_point(data = truenegs, aes(x = m, y = c(44, 46), color = condition)) + 
  geom_segment(data = truenegs, aes(x = cil, xend = cih, y = c(44, 46), yend = c(44, 46), color = condition)) + 
  scale_fill_grey("Condition") + scale_color_grey("Condition") +
  xlab("Response") + ylab("Count") +
  plot.style
```

### Statistical models

Basic continuous models:

Table 2 in paper: 
```{r}
contrasts(d.exp2$condition) <- rbind(.5, -.5)
contrasts(d.exp2$condition)

contrasts(d.exp2$truth) <- rbind(.5, -.5)
contrasts(d.exp2$truth)

#maximal model (truth * condition | item) does not converge
#This is the maximal model that converges
model.noage <- lmer(resp ~ condition * truth  
                    + (truth|subid) 
                    + (truth|item), 
                    data = filter(d.exp2, sent.type ==  "Negative"))
kable(summary(model.noage)$coefficients, digits = 2)
```

Including agegroup does not improve model fit:
```{r}
model.agegroup <- lmer(resp ~ condition * agegroup * truth  
                  + (truth|subid) 
                  + (truth|item), 
                  data = filter(d.exp2, sent.type ==  "Negative"))
kable(summary(model.agegroup)$coefficients, digits = 2)

anova(model.agegroup, model.noage)

```

Including age does not improve model fit:
```{r}
model.age <- lmer(resp ~ condition * age * truth  
                  + (truth|subid) 
                  + (truth|item), 
                  data = filter(d.exp2, sent.type ==  "Negative"))
kable(summary(model.age)$coefficients, digits = 2)

anova(model.age, model.noage)

```

Including age does not improve model fit:
```{r}
model.gender <- lmer(resp ~ condition * gender * truth  
                  + (truth|subid) 
                  + (truth|item), 
                  data = filter(d.exp2, sent.type ==  "Negative"))
kable(summary(model.gender)$coefficients, digits = 2)

anova(model.gender, model.noage)

```

What about just true negs?

```{r}
model.kids.trueneg <- lmer(resp ~ condition
                    + (1 |subid) 
                    + (condition |item), 
                    data = filter(d.exp2, sent.type ==  "Negative" & truth == "True"))
kable(summary(model.kids.trueneg)$coefficients, digits = 2)

```

T-tests:

```{r}
#All kids
subs <- aggregate(resp ~ subid + condition, d.exp2, mean)
t.test(resp ~ condition, subs)

#Three-year-olds
threes <- filter(d.exp2, agegroup ==  "3")
threes_subs <- aggregate(resp ~ subid + condition, threes, mean)
t.test(resp ~ condition, threes_subs)

#Four-year-olds
fours <- filter(d.exp2, agegroup ==  "4")
fours_subs <- aggregate(resp ~ subid + condition, fours, mean)
t.test(resp ~ condition, fours_subs)

```


